---
title: "reformat_nutrients_chla"
author: "Catarina Pien"
date: "1/20/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(readr)
library(lubridate)
library(tidyverse)
require(discretewq)
library(plotly)
```

Read in raw
```{r}
# EMP data, for reference on columns
emp <- wq(Sources = "EMP")

# nutrients and chlorophyll
nuts0 <- read_csv("data_raw/YBFMP_Chla_Nutrients.csv")
station_latlons <- read_csv("data_raw/Stations.csv") %>%
  rename(Station = StationCode)
```

Filter and clean nutrients
- stations of interest, fields of interest
- rename and remove variables
```{r}
stations_tokeep <- c("A0D82120386",  "B9D82851352", "A0200000")

nuts <- nuts0 %>%
  select(-DataOwner, -DataStatus, -LongStationName, -CASReg_Number, -Matrix, -Description, -ResultRejected, -Method, -Depth, -Units, -ParentSample) %>%
  filter(!is.na(CollectionDate),
         !(Analyte%in% c("Field Notes", "Field pH", "Field Weather Observations", "*No Lab Analyses (Field Measures Only)", "Field Water Temperature", "Field Secchi Depth", "Field Chlorophyll Volume", "Field Dissolved Oxygen", "Field Turbidity")))%>%
  filter(StationNumber %in% stations_tokeep) %>%
  mutate(CollectionDate = mdy_hm(CollectionDate),
         fMonth = factor(month(CollectionDate)),
         fYear = factor(year(CollectionDate))) %>%
  mutate(Sign = ifelse(grepl("<", Result), "<", "=")) %>%
  mutate(ShortStationName = case_when(StationNumber == "A0200000" ~ "SHR",
                             StationNumber == "B9D82851352" ~ "LIS",
                             StationNumber == "A0D82120386" ~ "STTD",
                             TRUE ~ ShortStationName)) %>%
  rename(Station = ShortStationName,
         Datetime = CollectionDate) %>%
  mutate(Index = 1:nrow(.)) %>%
  select(Index, everything())

unique(nuts$Analyte)
unique(nuts$Station)
```


For RL: 
* If no collection, value: NA, sign: =
* If collection and normal, value: numeric, sign: =
* If RL unknown, value: NA, sign: <
* If collection and <RL, value: numeric, sign: <
```{r}
# For now make all <RLs RL.
# What does <0 mean? <RL?
nuts_noRL <- nuts %>%
  mutate(Result2 = ifelse(Sign == "<", RptLimit, Result),
         Result2 = as.numeric(Result2),
         Year = year(Datetime),
         WY = ifelse(month(Datetime) > 9, Year + 1, Year),
         fWY = factor(WY))

str(nuts_noRL)
```


Plot data
```{r}
chl <- filter(nuts_noRL, Analyte %in% c("Chlorophyll a", "Pheophytin a"))
nonchl <- filter(nuts_noRL, !Analyte %in% c("Chlorophyll a", "Pheophytin a"))

# All data
plot_all <- ggplot(nuts_noRL, aes(x = Datetime, y = Result2, color = Analyte,
                                  text = paste("Index:", Index, "Datetime:", Datetime, "Station:", Station))) + geom_point(size = 0.5) + facet_wrap(~Analyte, scales = "free")
ggplotly(plot_all, tooltip = c("Datetime", "text"))

# Chl-a: 117, 191
# Pheophytin: 33, 38, 71
# Dissolved Calcium: 8806
# Dissolved Nitrate + Nitrite: 6476
# DOP: 4915
# TDS: 6301
# TSS: 11764 possibly but not that strong
# VSS: 11765
visual_outliers <- c(117, 191, 33, 38, 71, 8806, 6476, 4915, 6301, 11764, 11765)

# Chlorophyll only
monthbox <- ggplot(chl,aes(x = fMonth, y = Result2, fill = Analyte, text = paste("Index:", Index, "Datetime:", Datetime, "Station:", Station))) + geom_boxplot() + facet_wrap(Station~Analyte, scales = "free")
ggplotly(monthbox, tooltip = "text")

(yearbox <- ggplot(chl,aes(x = fWY, y = Result2, color = Analyte)) + geom_boxplot() + facet_wrap(Station~Analyte, scales = "free") + theme(axis.text.x = element_text(angle = 90)))
ggplotly(yearbox)

# Non-chlorophyll
ggplot(nonchl,aes(x = fMonth, y = Result2, color = Analyte)) + geom_boxplot() + facet_wrap(~Analyte, scales = "free")

ggplot(nonchl,aes(x = fMonth, y = Result2, color = Analyte)) + geom_boxplot() + facet_wrap(~Analyte, scales = "free")+ theme(axis.text.x = element_text(angle = 90))
```

Identify potential outliers
```{r}
library(ODWGtools)
outliers_test <- nuts_noRL %>% group_by(Station, Analyte) %>%
  mutate(MAD = ODWGtools::outlier_mad(Result2),
         Tukey = ODWGtools::outlier_tukey(Result2)) %>%
  ungroup() %>%
  mutate(Out_vis = ifelse(Index %in% c(visual_outliers), 1L, 0L))

outliers <- filter(outliers_test, MAD != "not outlier" & Tukey!= "not outlier") %>%
  select(Index, MAD, Tukey, Out_vis) %>%
  mutate(flag = ifelse(MAD == "extreme outlier" & Tukey == "extreme outlier" & Out_vis == 1, "3", ifelse((MAD == "extreme outlier" & Out_vis ==1) | (Tukey == "extreme outlier" & Out_vis == 1) | (MAD == "extreme outlier" & Tukey == "extreme outlier"), "2", "1")))
```

Visualize original flags
```{r}
outlier_vis <- left_join(nuts_noRL, outliers, by = "Index") %>%
  mutate(MAD = replace(MAD, is.na(MAD), "not outlier"),
         Tukey = replace(Tukey, is.na(Tukey), "not outlier"),
         Out_vis = replace(Out_vis, is.na(Out_vis), 0L)) 

# SHR
shr_outliers <- outlier_vis %>%
  filter(Station == "SHR") %>%
  ggplot() + 
  geom_point(aes(x = Datetime, y = Result2, color = flag, text = paste("Index:", Index)), size = 0.5) + facet_wrap(~Analyte, scales = "free")
ggplotly(shr_outliers)

 
# STTD & LIS
yolo_outliers <- outlier_vis %>%
  filter(Station != "SHR") %>%
  ggplot() + 
  geom_point(aes(x = Datetime, y = Result2, color = flag, text = paste("Index:", Index)), size = 0.5) + 
  facet_wrap(~Analyte, scales = "free")
ggplotly(yolo_outliers)

```


Modify flags based on visualized outliers
```{r}
shouldbe3 <- c(7985, 4935, 4935)
shouldbe1 <- c(2062, 2063, 11529, 7274, 8188, 1254, 1294, 1358, 10062,10126, 8895, 7217, 8253) 

outlier_mod <- outlier_vis %>%
  mutate(flag = ifelse(Index %in% shouldbe3, 3, ifelse(Index %in% shouldbe1 | (Station == "SHR" & Analyte == "Pheophytin a" & Result2 < 10) | (Station %in% c("STTD", "LIS") & Analyte == "Chlorophyll a" & Result2 < 100) | (Station %in% c("STTD", "LIS") & Analyte %in% c("Dissolved Ammonia", "Pheophytin a") & flag == 2), 1, flag))) %>%
  mutate(flag_comment = ifelse(!is.na(flag) & flag>1, Analyte, NA)) 

# are there samples that are flagged multiple times? 
check_comments <- outlier_mod %>%
  filter(!is.na(flag)) %>%
  group_by(SampleCode) %>%
  mutate(n = n()) %>%
  filter(n>1)
```

Plot visualized outliers
```{r}
# SHR
shr_outliers <- outlier_mod %>%
  filter(Station == "SHR") %>%
  ggplot() + 
  geom_point(aes(x = Datetime, y = Result2, color = flag, text = paste("Index:", Index)), size = 0.5) + facet_wrap(~Analyte, scales = "free")
ggplotly(shr_outliers)

 
# STTD & LIS
yolo_outliers <- outlier_mod %>%
  filter(Station != "SHR") %>%
  ggplot() + 
  geom_point(aes(x = Datetime, y = Result2, color = flag, text = paste("Index:", Index)), size = 0.5) + 
  facet_wrap(~Analyte, scales = "free")
ggplotly(yolo_outliers)
```


For now, ignore replicates/duplicates/blanks/spikes
```{r}
chla_nuts <- outlier_mod %>% 
  mutate(SampleType = ifelse(grepl("Duplicate", Notes), "Field Duplicate", SampleType)) %>%
  mutate(SampleType = ifelse(grepl("Matrix Spike", Notes), "Matrix Spike", SampleType)) %>%
  filter(SampleType == "Normal Sample") 

```

check sample numbers, don't want there to be duplicates by datetime/analyte
- Don't know why there are still duplicates. Probably an unlabeled dup/rep?
- Pick just the first one of each datetime (slice). 
```{r}
samples <- chla_nuts %>% group_by(Station, Datetime, Analyte) %>%
  mutate(n = n()) %>%
  filter(n>1)

chla_nuts_slice <- chla_nuts %>%
  group_by(Station, Datetime, Analyte) %>%
  slice(1) %>%
  ungroup %>%
  select(-SampleType, -Result, -RptLimit)
```


```{r}
chla_nuts_wide_test <- chla_nuts_all %>% 
  mutate(flag_name = paste0("flag_", Analyte),
         sign_name = paste0("sign_", Analyte),
         flag_label = "flag")%>%
  pivot_wider(
    names_from = c(Analyte),
    id_cols = c(SampleCode, Station, StationNumber, fWY, Datetime, fMonth),
    values_from = c(Result2, flag, Sign)) 
```

change to wide-v1 (all included)
```{r}
chla_nuts_all <- select(chla_nuts_slice, -flag_comment, -MAD, -Tukey, -Out_vis, -Index)
chla_nuts_wide <- chla_nuts_all %>% pivot_wider(names_from = (Analyte), values_from = c(Result2, Sign))
str(chla_nuts_wide)
```


change to wide -v2 (remove flags of 3)
```{r}
chla_nuts_filt <- chla_nuts_slice %>%
  filter(is.na(flag) | flag <3 ) %>%
  select(-flag, -flag_comment, -MAD, -Tukey, -Out_vis, -Index)
chla_nuts_wide_filt <- chla_nuts_filt %>% pivot_wider(names_from = (Analyte), values_from = c(Result2, Sign))
```


Rename columns again
```{r}
chla_nuts_clean <- chla_nuts_wide %>%
  left_join(station_latlons, by = "Station") %>%
  mutate(Source = "YBFMP") %>%
  select(Station, Latitude, Longitude, Datetime, Notes, SampleCode,
         Chlorophyll = `Result2_Chlorophyll a`,
         Pheophytin = `Result2_Pheophytin a`,
         DissAmmonia_Sign = `Sign_Dissolved Ammonia`,
         DissAmmonia = `Result2_Dissolved Ammonia`,
         DissCalcium_Sign = `Sign_Dissolved Calcium`,
         DissCalcium = `Result2_Dissolved Calcium`,
         DissChloride_Sign = `Sign_Dissolved Chloride`,
         DissChloride = `Result2_Dissolved Chloride`,
         DissNitrateNitrite_Sign = `Sign_Dissolved Nitrate + Nitrite`,
         DissNitrateNitrite = `Result2_Dissolved Nitrate + Nitrite`,
         DOC_Sign = `Sign_Dissolved Organic Carbon`,
         DOC = `Result2_Dissolved Organic Carbon`,
         TOC_Sign = `Sign_Total Organic Carbon`,
         TOC = `Result2_Total Organic Carbon`,
         DON_Sign = `Sign_Dissolved Organic Nitrogen`,
         DON = `Result2_Dissolved Organic Nitrogen`,
         DissOrthophos_Sign = `Sign_Dissolved ortho-Phosphate`,
         DissOrthophos = `Result2_Dissolved ortho-Phosphate`,
         TotPhos_Sign = `Sign_Total Phosphorus`,
         TotPhos = `Result2_Total Phosphorus`,
         DissSilica_Sign = `Sign_Dissolved Silica (SiO2)`,
         DissSilica = `Result2_Dissolved Silica (SiO2)`,
         TDS_Sign = `Sign_Total Dissolved Solids`,
         TDS = `Result2_Total Dissolved Solids`,
         TSS_Sign = `Sign_Total Suspended Solids`,
         TSS = `Result2_Total Suspended Solids`,
         VSS_Sign = `Sign_Volatile Suspended Solids`,
         VSS = `Result2_Volatile Suspended Solids`,
         TKN_Sign = `Sign_Dissolved Total Kjeldahl Nitrogen`,
         TKN = `Result2_Total Kjeldahl Nitrogen`,
         Year,
         WY,
         Month = fMonth)
```

Rename columns for flag-filtered dataset
```{r}
chla_nuts_clean_filt <- chla_nuts_wide_filt %>%
  left_join(station_latlons, by = "Station") %>%
  mutate(Source = "YBFMP") %>%
  select(Station, Latitude, Longitude, Datetime, Notes, SampleCode,
         Chlorophyll = `Result2_Chlorophyll a`,
         Pheophytin = `Result2_Pheophytin a`,
         DissAmmonia_Sign = `Sign_Dissolved Ammonia`,
         DissAmmonia = `Result2_Dissolved Ammonia`,
         DissCalcium_Sign = `Sign_Dissolved Calcium`,
         DissCalcium = `Result2_Dissolved Calcium`,
         DissChloride_Sign = `Sign_Dissolved Chloride`,
         DissChloride = `Result2_Dissolved Chloride`,
         DissNitrateNitrite_Sign = `Sign_Dissolved Nitrate + Nitrite`,
         DissNitrateNitrite = `Result2_Dissolved Nitrate + Nitrite`,
         DOC_Sign = `Sign_Dissolved Organic Carbon`,
         DOC = `Result2_Dissolved Organic Carbon`,
         TOC_Sign = `Sign_Total Organic Carbon`,
         TOC = `Result2_Total Organic Carbon`,
         DON_Sign = `Sign_Dissolved Organic Nitrogen`,
         DON = `Result2_Dissolved Organic Nitrogen`,
         DissOrthophos_Sign = `Sign_Dissolved ortho-Phosphate`,
         DissOrthophos = `Result2_Dissolved ortho-Phosphate`,
         TotPhos_Sign = `Sign_Total Phosphorus`,
         TotPhos = `Result2_Total Phosphorus`,
         DissSilica_Sign = `Sign_Dissolved Silica (SiO2)`,
         DissSilica = `Result2_Dissolved Silica (SiO2)`,
         TDS_Sign = `Sign_Total Dissolved Solids`,
         TDS = `Result2_Total Dissolved Solids`,
         TSS_Sign = `Sign_Total Suspended Solids`,
         TSS = `Result2_Total Suspended Solids`,
         VSS_Sign = `Sign_Volatile Suspended Solids`,
         VSS = `Result2_Volatile Suspended Solids`,
         TKN_Sign = `Sign_Dissolved Total Kjeldahl Nitrogen`,
         TKN = `Result2_Total Kjeldahl Nitrogen`,
         Year,
         WY,
         Month = fMonth)
```

Select down to columns that are in Sarah's dataset
```{r}
chla_nuts_publish <- select(chla_nuts_clean, -DissCalcium_Sign, -DissChloride_Sign, -DOC_Sign, -TOC_Sign, -DON_Sign, -TotPhos_Sign, -DissSilica_Sign, -TDS_Sign, -TSS_Sign, -VSS_Sign, -TKN_Sign)

# Will later need to add Source, MonthYear, StationID (YBFMP + Station), Season
```

Same, but for flag-filtered version
```{r}
chla_nuts_filt_publish <- select(chla_nuts_clean_filt, -DissCalcium_Sign, -DissChloride_Sign, -DOC_Sign, -TOC_Sign, -DON_Sign, -TotPhos_Sign, -DissSilica_Sign, -TDS_Sign, -TSS_Sign, -VSS_Sign, -TKN_Sign)
```

Write cleaned up file to be used later (merged with zooplankton, etc.)
```{r, eval = FALSE}
today = format(today(),"%Y%m%d")
write.csv(chla_nuts_publish, paste0("data_clean/clean_ybfmp_chla_nuts_", today, ".csv"), row.names = FALSE)
```

Write filtered version
```{r}
write.csv(chla_nuts_filt_publish, paste0("data_clean/clean_ybfmp_chla_nuts_filt_", today, ".csv"), row.names = FALSE)
```

