---
title: "clean_phys_wq"
author: "Catarina Pien"
date: "1/25/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setup
```{r}
library(tidyverse)
library(knitr)
library(lubridate)
library(kableExtra)
library(gridExtra)
library(readr)
library(readxl)
library(plotly)
```

##2. Read in wq
```{r}
WDL <- read_csv("data_raw/WDL_SAM_Codes_20201002.csv")
#station_latlons <- read_csv("data_raw/Stations.csv") %>%
#  rename(Station = StationCode)
phys0 <- read_csv("data_raw/TblPhysicalData_20210204.csv")
physrecent <- read_excel("data_raw/YBFMP_WQ_Data_WORKING_20200721.xlsx", skip=1)
light <- read_csv("data_raw/TblLight.csv")
station_simple <- read.csv("data_raw/YBFMP_Stations_Coordinates_Info_20200716.csv") %>%
  select(1:2) %>% 
  rename(StationCode = Station.Name, StationNumber = Station.Number)
source("WQ_plots.R")
```

```{r}
corrplots <- function(df, x, y) {
  ggplot(df, aes(x={{x}}, y = {{y}})) + 
    geom_point() + 
    geom_smooth() +
  #labs(title = paste0(x, " by " y), x = x, y = y)+
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 11))
}
```

## 3. Rename and restructure variables

Datetime
```{r rename}
phys <- phys0 %>% mutate(Datetime = paste(Date, Time, sep = " "),
                        Datetime = mdy_hms(Datetime),
                        Date = date(Datetime),
                        Year = year(Datetime),
                        Month = month(Datetime),
                        #MonthAbb = month(Datetime, label = TRUE), 
                        Tide = as.factor(Tide))
```

Physical Variables
```{r}
phys.s <- phys %>%
  select(PhysicalDataID, StationCode = `Station Code`,
         Date, Datetime, Year, Month,  Tide,
         Microcystis = MicrocystisVisualRank,
         Secchi = SecchiDiskDepth, WaterTemp = WaterTemperature,
         Conductivity = EC, SpecificConductance = SpCnd,
         DO, pH, Turbidity, 
         FieldComments_WQ = Comments)
```



Light table - not currently included, but might want to include this? 
```{r}
light.s <- light %>%
  rename(FieldComments_Light = FieldComments) %>%
  select(-Field1)
```

WDL (to later join with phys)
- select down to certain columns. Then remove certain columns and rename. 
```{r}
unique(WDL$SPP_DESCRIPTION)
WDL.s <- WDL %>% 
  select(c(RUN_NAME, ACT_UNIT_NAME, STATION_NUM, NAME, SAM_CODE, SAM_COLLECTION_DATE, SAM_SAM_CODE_PARENT, SPP_DESCRIPTION)) %>%
  filter(RUN_NAME == "YB - Fall Study" | RUN_NAME == "YB - Fall Study - SFSU 2 Depths" | ACT_UNIT_NAME == "Aquatic Ecology") %>%
  select(-RUN_NAME, -ACT_UNIT_NAME) %>%
  rename(
    StationNumber = STATION_NUM,
    StationName = NAME, 
    SAM_Code = SAM_CODE,
    SAM_Code_Parent = SAM_SAM_CODE_PARENT,
    Datetime = SAM_COLLECTION_DATE,
    SampleType = SPP_DESCRIPTION) %>%
  filter(SampleType %in% c("Normal Sample", "Duplicate Sample"))
  

WDL.s$Datetime <- mdy_hm(WDL.s$Datetime)
WDL.s$SamplingDate <- as.Date(WDL.s$Datetime)
```

### Format more recent data
```{r}
physK <- physrecent[-1,] %>%
  janitor::clean_names(case = "upper_camel") %>%
  select(-StationNumber) %>%
  rename(StationCode = StationName) %>%
  filter(StationCode %in% c("LIS", "SHR", "STTD")) %>%
  left_join(station_simple, by = "StationCode") %>%
  mutate(PhysicalDataID = NA,
         WdlSamCollectionDate = as.Date(as.numeric(as.character(WdlSamCollectionDate)), origin = "1899-12-30"),
          Time = as.numeric(CollectionTime),
         Hour = floor(Time * 24),
         Minute = (((Time*24) - Hour) * 60),
         Time2 = paste0(Hour, ":", Minute),
         time = strptime(Time2, format="%H:%M")  %>%
           strftime(Time2, format = "%H:%M"),
         Datetime = paste(WdlSamCollectionDate, time),
        Datetime = ymd_hm(Datetime),
         Year = year(Datetime),
         Month=month(Datetime),
        Microcyst = as.numeric(Microcyst)) %>%
  select(PhysicalDataID, 
         StationCode,
         SamplingDate = WdlSamCollectionDate,
         Datetime,
         Year,
         Month,
         Tide,
         Microcystis = Microcyst,
         Secchi,
         WaterTemp,
         Conductivity = Ec,
         SpecificConductance = SpCond,
         DO = DoProbe,
         pH = PH,
         Turbidity = Turb,
         FieldComments_WQ = FieldCommentsWq,
         StationNumber,
         SAM_Code = SamplingNumberEvent) %>%
  mutate(across(.cols = Secchi:Turbidity, as.numeric))
```

## 4.Filter to relevant stations and timeframe
```{r filter, results = FALSE, message = FALSE, warning = FALSE}
# Filter only SHR and STTD
unique(phys.s$StationCode)
phys_yolo <- phys.s %>%
  filter(StationCode %in% c("LIS", "SHR", "STTD"))
```

## 5.Look at observations in stations
```{r Stations, message = FALSE, warning = FALSE}

# summarize number of observations per year
sta.sum <- phys_yolo %>%
  group_by(StationCode, Year) %>%
  summarize(n = n())

# plot number of observations per station
ggplot(sta.sum, aes(Year, n, fill = StationCode)) + geom_bar(stat = "identity") +theme_bw() +
  scale_y_continuous(breaks = seq(0,300,20))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 11))

```


## 6.Summary Tables - Min and max of variables by station

```{r minmaxtables, message = FALSE, warning = FALSE}
# All database data
WQ.overall <- phys_yolo %>%
  group_by(StationCode) %>%
  summarize(min.temp = min(WaterTemp, na.rm=T),
            max.temp = max(WaterTemp, na.rm=T),
            min.Conductivity = min(Conductivity, na.rm=T),
            max.Conductivity = max(Conductivity,na.rm=T),
            min.SPC = min(SpecificConductance,na.rm=T),
            max.SPC = max(SpecificConductance,na.rm=T),
            min.Secchi = min(Secchi,na.rm=T),
            max.Secchi = max(Secchi,na.rm=T),
            min.Turbidity = min(Turbidity,na.rm=T),
            max.Turbidity = max(Turbidity,na.rm=T),
            min.pH = min(pH,na.rm=T),
            max.pH = max(pH,na.rm=T),
            min.DO = min(DO,na.rm=T),
            max.DO = max(DO,na.rm=T),
            n = n()) 

kable(t(WQ.overall)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

## 7. Dot plots
```{r}
WQ_long <- pivot_longer(phys_yolo, cols = c(WaterTemp:Turbidity),names_to = "Parameter", values_to = "Value") %>%
  mutate(Index = 1:nrow(.))
```

```{r WQ Plot, message = FALSE, warning = FALSE}
WQ_plots <- WQ_long %>%
  ggplot(aes(x= Datetime, y = Value, color = StationCode, text = paste("Index:", Index, "PhysicalDataID:", PhysicalDataID))) + geom_point(size = 0.5) +  facet_wrap(~Parameter, scales = "free")
ggplotly(WQ_plots)
```


##8. More Plots

```{r Yolo temperature plots, warning = FALSE, message = FALSE}
Yearbox(phys_yolo, WaterTemp)
Monthbox(phys_yolo,WaterTemp)
VisHist(phys_yolo, WaterTemp, 5)

Yearbox(phys_yolo, Secchi)
Monthbox(phys_yolo,Secchi)
VisHist(phys_yolo, Secchi, 0.5)

Yearbox(phys_yolo, Turbidity)
Monthbox(phys_yolo,Turbidity)
VisHist(phys_yolo, Turbidity, 20)

Yearbox(phys_yolo, DO)
Monthbox(phys_yolo,DO)
VisHist(phys_yolo, DO, 1)

Yearbox(phys_yolo, Conductivity)
Monthbox(phys_yolo,Conductivity)
VisHist(phys_yolo, Conductivity, 20)

Yearbox(phys_yolo, SpCnd)
Monthbox(phys_yolo,SpCnd)
VisHist(phys_yolo, SpCnd, 20)

Yearbox(phys_yolo, pH)
Monthbox(phys_yolo,pH)
VisHist(phys_yolo, pH, 1)

```


##9. Correlation Plots: TurbidityxSecchi, DOxTemp, ConductivityxSpCnd
```{r Correlations, message = FALSE, warning = FALSE}
sec.tur <- corrplots(phys_yolo, Secchi, Turbidity)
do.temp <- corrplots(phys_yolo, WaterTemp, DO)
ec.spc <- corrplots(phys_yolo, Conductivity, SpecificConductance)
grid.arrange(sec.tur, do.temp, ec.spc)
```



## 10. QC data

Outlier checking - here is some code I have been considering for our fish data, but I am not currently using this to make any decisions on flagging.
```{r}
library(ODWGtools)
Outliers0 <- WQ_long %>%
  group_by(StationCode, Parameter) %>%
  arrange(StationCode, Datetime, Parameter) %>%
  mutate(Tukey = outlier_tukey(Value, na.rm = TRUE),
         MAD = outlier_mad(Value, mask = !is.na(Value)),
         Tukey_num = ifelse(Tukey == "extreme outlier", 2L, ifelse(Tukey == "mild outlier", 1L, ifelse(Tukey == "not outlier", 0L, Tukey))),
         MAD_num = ifelse(MAD == "extreme outlier", 2L, ifelse(MAD == "mild outlier", 1L, ifelse(MAD == "not outlier", 0L, MAD))),
         lag = difftime(Datetime, lag(Datetime), units = "days"))%>%
  ungroup() 
```


Visual identification checks 
```{r QC, warning = FALSE, message = FALSE}
# Edits to data
to_drop <- c(1575, 1631, 1632, 1722, 1724, 1867, 1308, 748)

# To change
phys_yolo$WaterTemp[phys_yolo$PhysicalDataID == 688] <- 16.7 # Was Fahrenheit
phys_yolo$WaterTemp[phys_yolo$WaterTemp == 0.169] <- 16.9 # Decimal place
phys_yolo$Turbidity[phys_yolo$PhysicalDataID==680] <- 15.4 # Typo
phys_yolo$Turbidity[phys_yolo$PhysicalDataID==656] <- 94.2 # Typo
phys_yolo$Conductivity[phys_yolo$PhysicalDataID==748] <- 732 # Typo
phys_yolo$Conductivity[phys_yolo$PhysicalDataID==926] <- 1284 # Typo
phys_yolo$Conductivity[phys_yolo$PhysicalDataID==1847] <- 170 # Typo
phys_yolo$SpecificConductance[phys_yolo$PhysicalDataID==1140] <- 848 # Typo
phys_yolo$SpecificConductance[phys_yolo$PhysicalDataID==1139] <- 889 # Typo
phys_yolo$SpecificConductance[phys_yolo$PhysicalDataID==1847] <- 101 # Typo
phys_yolo$Secchi[phys_yolo$PhysicalDataID==708] <- NA # Previously 0, but actually was not taken
phys_yolo$Secchi[phys_yolo$PhysicalDataID==708] <- NA # Previously 0, but actually was not taken
phys_yolo$DO[phys_yolo$PhysicalDataID==1409] <- NA # Previously 0, but actually was not taken
phys_yolo$pH[phys_yolo$PhysicalDataID==1501] <- 7.92 # Typo


# One entry was not entered
phys_yolo$StationCode[phys_yolo$PhysicalDataID==1752] <- "STTD"
phys_yolo$Conductivity[phys_yolo$PhysicalDataID==1752] <- 530
phys_yolo$SpecificConductance[phys_yolo$PhysicalDataID==1752] <- 563
phys_yolo$Turbidity[phys_yolo$PhysicalDataID==1752] <- 16.5
phys_yolo$Tide[phys_yolo$PhysicalDataID==1752] <- "Ebb"
phys_yolo$Date[phys_yolo$PhysicalDataID==1752] <-as.Date("2018-09-27", format = "%Y-%m-%d")
phys_yolo$pH[phys_yolo$PhysicalDataID==1752] <- 8.13

# Formatting
# phys_yolo$Time[phys_yolo$PhysicalDataID==401] <- hms("13:22:00")
# phys_yolo$Time[phys_yolo$PhysicalDataID==1602] <- hms("13:50:00")

# Remove rows with no data
phys_yolo <- phys_yolo%>%
  filter(!PhysicalDataID %in% to_drop)
```


## 11. Flag Data
* Flag data 
* 3: Replaced (above changes, mostly typos)
* 2: Suspect data, based on plots, but not out of the realm of possibility, not confirmed in field data
* blank: Pass
* Include comment to say which variable is flagged/replaced.
```{r Flag, message = FALSE, warning = FALSE}
# Lists for flags and flag comments
# Modified data due to typos
modified <- c(688, 680, 656, 748, 926, 1847, 1140, 1139, 1847, 708, 1409, 1501, 1818)

# Flagged but not replaced
flag <-  c(899, 885, 1087, 1090, 1314, 1218, 602, 668, 841, 1187)

# Variable associated with flag
wtemp <- c(688, 1818)
turb <- c(680, 656, 885, 668, 841, 1187)
cond <- c(748, 926, 1847)
spc <- c(1140, 1139, 1847, 899, 1087)
sec <- c(708, 1090)
do <- c(1409, 1314, 1218)
ph <- c(1501)

# Adding Flag and Comment variables
phys_yolo_flag <- phys_yolo %>%
  mutate(Flag_PQC = ifelse(PhysicalDataID %in% modified, 3,
                           ifelse(PhysicalDataID %in% flag, 2, "")),
         Comment_PQC = ifelse(PhysicalDataID %in% wtemp, "WTEMP",
                              ifelse(PhysicalDataID %in% turb, "TURB",
                                     ifelse(PhysicalDataID %in% cond, "COND",
                                            ifelse(PhysicalDataID %in% spc, "SPC",
                                                   ifelse(PhysicalDataID %in% sec, "SECCHI",
                                                          ifelse(PhysicalDataID %in% do, "DO",
                                                          ifelse(PhysicalDataID %in% ph, "PH", ""))))))))

phys_yolo_flag$Flag_PQC <- as.factor(phys_yolo_flag$Flag_PQC)

```

## 12. Plot QCed data

```{r Plot edited data, message = FALSE, warning = FALSE, echo = FALSE}
colors = c("lightslategrey", "goldenrod2", "darksalmon")

phys_yolo_flag_long <- pivot_longer(phys_yolo_flag, cols = c(WaterTemp:Turbidity),names_to = "Parameter", values_to = "Value") %>%
  mutate(Index = 1:nrow(.))

WQ_outliers <- phys_yolo_flag_long %>%
ggplot() + 
  geom_point(aes(Datetime, Value, colour = factor(Flag_PQC), text = PhysicalDataID), size = 1) + 
  facet_wrap(~Parameter, scales = "free") + 
  scale_colour_manual(values = c("lightslategrey", "goldenrod2", "darksalmon")) + 
  theme_bw() 
WQ_outliers
ggplotly(WQ_outliers)
unique(phys_yolo_flag_long$Flag_PQC)

# p.temp <- ggplot(phys_yolo_flag, aes(Date, WaterTemp, col = Flag_PQC)) + geom_point() + labs(y = "Water Temp(C)") + scale_colour_manual(values = cols)+
#   theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.secchi <- ggplot(phys_yolo_flag, aes(Date, Secchi, col = Flag_PQC)) + geom_point() + labs(y = "Secchi Depth(m)") +
#  scale_colour_manual(values = cols)+ theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.conductivity <- ggplot(phys_yolo_flag, aes(Date, Conductivity, col = Flag_PQC)) + geom_point() + labs(y = "Conductivity(useimens/cm)") +
#   theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.spc <- ggplot(phys_yolo_flag, aes(Date, SpecificConductance, col = Flag_PQC)) + geom_point() + labs(y = "SpCnd(useimens/cm)") +
#   scale_colour_manual(values = cols)+theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.ec <- ggplot(phys_yolo_flag, aes(Date, Conductivity, col = Flag_PQC)) + geom_point() + labs(y = "EC(useimens/cm)") +
#   scale_colour_manual(values = cols)+theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.pH <- ggplot(phys_yolo_flag, aes(Date, pH, col = Flag_PQC)) + geom_point() + labs(y = "pH") +
#   scale_colour_manual(values = cols)+theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.do <- ggplot(phys_yolo_flag, aes(Date, DO, col = Flag_PQC)) + geom_point() + labs(y = "DO(mg/L)") +
#   scale_colour_manual(values = cols)+theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# p.turb <- ggplot(phys_yolo_flag, aes(Date, Turbidity, col = Flag_PQC)) + geom_point() + labs(y = "Turbidity(NTU)") +
#  scale_colour_manual(values = cols)+ theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 11))
# 
# # Arrange plots together
# grid.arrange(p.temp, p.ec, p.spc, p.secchi, p.turb, p.do, p.pH, ncol = 2)


```


## 13. Add SAM code
```{r}
# Join station
phys_sta <- left_join(phys_yolo_flag, station_simple, by = "StationCode")  %>%
  rename(SamplingDate = Date) 

# Join with WDL data. Remove duplicates.
WDL_phys <- left_join(phys_sta, WDL.s, by = c("StationNumber", "Datetime",  "SamplingDate")) %>%
  arrange(Datetime, StationCode) %>%
  filter(SampleType %in% c(NA, "Normal Sample"))

# Check for multiple entries for given station-datetime
sampsize <- WDL_phys %>%
  group_by(StationNumber, Datetime) %>%
  mutate(n = n()) %>%
  filter(n>1)

# Remove the second one - they are all the same!
WDL_phys_slice <- WDL_phys %>%
  group_by(StationNumber, Datetime) %>%
  mutate(n = n()) %>%
  slice(1)%>%
  select(-StationName, -SAM_Code_Parent, -SampleType, -n) %>%
  ungroup()
```


## 14. New data QAQC

```{r}
WQ_longK <- pivot_longer(physK, cols = c(Secchi:Turbidity),names_to = "Parameter", values_to = "Value") %>%
  mutate(Index = 1:nrow(.))
```

```{r WQ Plot, message = FALSE, warning = FALSE}
WQ_plotsK <- WQ_longK %>%
  ggplot(aes(x= Datetime, y = Value, color = StationCode, text = paste("Index:", Index, "PhysicalDataID:", PhysicalDataID))) + geom_point(size = 0.5) +  facet_wrap(~Parameter, scales = "free")
ggplotly(WQ_plotsK)
```


```{r Yolo temperature plots, warning = FALSE, message = FALSE}
Yearbox(physK, WaterTemp)
Monthbox(physK,WaterTemp)
VisHist(physK, WaterTemp, 5)

Yearbox(physK, Secchi)
Monthbox(physK,Secchi)
VisHist(physK, Secchi, 0.5)

Yearbox(physK, Turbidity)
Monthbox(physK,Turbidity)
VisHist(physK, Turbidity, 20)

Yearbox(physK, DO)
Monthbox(physK,DO)
VisHist(physK, DO, 1)

Yearbox(physK, Conductivity)
Monthbox(physK,Conductivity)
VisHist(physK, Conductivity, 20)

Yearbox(physK, SpecificConductance)
Monthbox(physK,SpecificConductance)
VisHist(physK, SpecificConductance, 20)

Yearbox(physK, pH)
Monthbox(physK,pH)
VisHist(physK, pH, 1)

```

```{r}
sec.tur2 <- corrplots(physK, Secchi, Turbidity)
do.temp2 <- corrplots(physK, WaterTemp, DO)
ec.spc2 <- corrplots(physK, Conductivity, SpecificConductance)
grid.arrange(sec.tur2, do.temp2, ec.spc2)
```

```{r}
library(ODWGtools)
OutliersNew <- WQ_longK %>%
  group_by(StationCode, Parameter) %>%
  arrange(StationCode, Datetime, Parameter) %>%
  mutate(Tukey = outlier_tukey(Value, na.rm = TRUE),
         MAD = outlier_mad(Value, mask = !is.na(Value)),
         Tukey_num = ifelse(Tukey == "extreme outlier", 2L, ifelse(Tukey == "mild outlier", 1L, ifelse(Tukey == "not outlier", 0L, Tukey))),
         MAD_num = ifelse(MAD == "extreme outlier", 2L, ifelse(MAD == "mild outlier", 1L, ifelse(MAD == "not outlier", 0L, MAD))),
         lag = difftime(Datetime, lag(Datetime), units = "days"))%>%
  ungroup() 

```

```{r}
WQ_outliersNewT <- OutliersNew %>%
ggplot() + 
  geom_point(aes(Datetime, Value, colour = factor(Tukey_num), text = PhysicalDataID), size = 1) + 
  facet_wrap(~Parameter, scales = "free") + 
  scale_colour_manual(values = c("lightslategrey", "goldenrod2", "darksalmon")) + 
  theme_bw() 
WQ_outliersNewT
ggplotly(WQ_outliersNewT)


WQ_outliersNewM <- OutliersNew %>%
ggplot() + 
  geom_point(aes(Datetime, Value, colour = factor(MAD_num), text = PhysicalDataID), size = 1) + 
  facet_wrap(~Parameter, scales = "free") + 
  scale_colour_manual(values = c("lightslategrey", "goldenrod2", "darksalmon")) + 
  theme_bw() 
WQ_outliersNewM
ggplotly(WQ_outliersNewM)
```


Some work needs to be done here to check for outliers for recent data. There is a EC/SPC that looks a bit weird, and some of the turbidities. 


## 15. Combine WQ files

```{r}
alldata <- bind_rows(WDL_phys_slice, physK)
```


## 16. Select final columns
```{r}
WQ_final <- alldata %>%
  select(SampleCode = SAM_Code, PhysicalDataID, StationCode, StationNumber, Datetime, Date = SamplingDate, Year, Month, Tide, Microcystis, Secchi, WaterTemp, Conductivity, SpecificConductance, DO, pH, Turbidity, Flag_PQC, Comment_PQC)
```


## 17. Replace blanks
```{r Write file, message = FALSE, warning = FALSE}
```


## 18. Write file
```{r, eval = FALSE}
# Write cleaned up file to be used later (merged with zooplankton, etc.)
today = format(today(),"%Y%m%d")
write.csv(WQ_final, paste0("data_clean/clean_WQ_phys_", today, ".csv"), row.names = FALSE)

```

